A spiking neural integrator model of the adaptive control of action by the medial prefrontal cortex

### 2013-12-05 paper/jneurosci2013.tex
Removed figures and footnote from paper
In order to conform to JNeuroscience submission guidelines

\newcommand{\scalefigone}[3]{
  \begin{figure}[ht!]
    % Requires \usepackage{graphicx}
    \centering
    \includegraphics[width=#2\columnwidth]{#1}
    \caption{#3}
    \label{#1}
  \end{figure}}

@@

\textbf{Pages:} 32. \textbf{Figures:} 10. \textbf{Tables:} 2. \textbf{Multimedia, 3D models:} 0

@@

The experimental data analyzed in this study is the same
as that in \citetext{Narayanan2009},
where more detailed explanation of the experimental
methods can be found.
In that study, twelve male Long-Evans rats
were trained to press down a lever
for a fixed amount of time (the ``foreperiod'';
1 second for this study)
using standard methods \cite{Laubach2000}.
After the foreperiod, an auditory cue is presented.
The subject then has 0.6 seconds (the response window)
in which to release the lever.
If the lever is released in that time,
the trial is a success, and the subject is rewarded
with water.
If the lever is released during the foreperiod,
the trial is classified as a ``premature error,''
and is penalized by a time-out with the house lights extinguished.
If the lever is not released within the response window,
the trial is classified as a ``late error,''
and is also penalized by a time-out with the house lights extinguished.
Figure~\ref{fig1}A presents a schematic view of the task.

@@

In order to provide appropriate
environmental input for
the networks described
in subsequent sections,
the same reaction time task is simulated
using a finite state machine.
The task can be in one of seven states~($S$):
trial start~(TS), foreperiod~(FP),
timeout~(TO), cue~(C), response window~(R),
reward~(Rw) and intertrial interval~(ITI).
State transitions can be seen in
the graphical depiction of
the finite state machine in Figure~\ref{fig1}C.

@@

In \citetext{Singh2006}, neural recordings
were analyzed during a single trial
of the vibrotactile discrimination task.
This corresponds to a single trial
of the simple RT task
in which the input, $u$,
is the lever press,
and the initial state, $x$, is $(0, 0)$.
However, in the simple RT task we perform
many consecutive trials.
We therefore use other inputs from
the environment to match
the trajectories tracked by
the experimental principal components
(shown in Figure~\ref{fig2}).

Since the principal components
resemble the double integrator network
in the post-correct case,
we aim to reset the system to the origin point
after correct trials.
Figure~\ref{fig2}D indicates that
the principal components return to their
original state after approximately 2 seconds,
which is the time at which
the subject receives reward.
Therefore, we hypothesize that
the trial's outcome (i.e., reward delivery
or a timeout) resets the system
to some particular state,
as proposed in \citetext{Narayanan2009}.
In the case of reward delivery,
we implement this hypothesis
by using the reward delivery signal, $u_{Rw}$,
to drive the system toward $(0, 0)$.
\begin{align*}
  \dot{x}_1 &= u - R \, u_{Rw} \, x_1 \\
  \dot{x}_2 &= \beta x_1 - R \, u_{Rw} \, x_2,
\end{align*}
where $R$ is a scalar parameter
indicating how strongly the system
is driven to the origin point.

In the post-error case,
the principal components
in Figure~\ref{fig2}D are
not at the origin point.
Since the lever press increases $x_1$,
when $x_1 > 0$ we can conclude that
the subject is currently performing the task.
We therefore hypothesize that $x_1 < 0$
indicates that an error has recently occurred.
We implement this hypothesis
by using the house lights signal,
$u_{TO}$, to decrease $x_1$.
\begin{equation} \label{eq:dint}
  \begin{aligned}
    \dot{x}_1 &= u - R \, u_{Rw} \, x_1 - E u_{TO} \\
    \dot{x}_2 &= \beta x_1 - R \, u_{Rw} \, x_2,
  \end{aligned}
\end{equation}
where $E$ is a scalar parameter
controlling how quickly $x_1$ decreases.
Note that $x_2$ is not directly driven by $u_{TO}$;
however, because of the $\beta x_1$ term,
$x_2$ will also decrease.

Finally, Figure~\ref{fig2}D shows that
the principal components are unstable
during the intertrial interval.
To model trial-by-trial cycles of
excitability associated with task engagement,
a slow oscillation ($\sim0.2$ Hz)
was applied to the networks.
The final dynamics of the double integrator are
\begin{equation} \label{eq:dint-o}
  \begin{aligned}
    \dot{x}_1 &= u - R \, u_{Rw} \, x_1 - E u_{TO} + \sin(0.2 \cdot 2 \pi) \\
    \dot{x}_2 &= \beta x_1 - R \, u_{Rw} \, x_2,
  \end{aligned}
\end{equation}
where $\beta$, $R$, and $E$ are positive scalars.
The effects of changing these free parameters
are explored in Section~\ref{sec:params}.

@@

The previous sections describe
a dynamical system hypothesis
of the mPFC and surrounding areas
of cortex that together
perform the simple reaction time task
adaptively depending on the outcome
of the previous trial.
While these can be numerically simulated
to observe ideal system behavior,
we construct a spiking neural network
implementing the above equations
in order to test the hypothesis
in a biologically plausible setting
that approximates the direct numerical simulation.
Testing this neural implementation
allows us to determine if the approximation
is able to perform the task
in a manner comparable to the real brain
by producing simulated neural data
that can be compared directly with
experimental neural data.
We construct spiking neural networks
using the principles of the Neural Engineering Framework
(NEF; \citenoparens{Eliasmith2003})
in the Nengo simulation environment
\cite{BekolayInpress}.\footnote{The Nengo scripts and simulation
  package simulating the networks and their environment are available
  at \url{https://github.com/tbekolay/jneurosci2013/releases/tag/2013-11-29}.}
Using the NEF, each signal
from the dynamical systems above
are represented with a population of spiking neurons,
and transformed through the connections
between those populations
(see Figure~\ref{fig3}).

@@

Simulated spike trains were generated
by the Nengo simulator.
The populations representing signals
in the double integrator and control
networks (see Sections~\ref{sec:dint} and \ref{sec:control})
contained 1200 adaptive leaky integrate-and-fire neurons
(see Figure~\ref{fig3});
174 of the neurons in an analyzed population
were randomly selected
in order to match the number of units
from the experimental study.

@@

As previously described in \citetext{Narayanan2009}
and summarized in Figure~\ref{fig2},
the first two principal components
of the experimental data
are modulated around the lever press
and appear to track
whether the subject has pressed the lever and
the relative amount of time
that the lever has been pressed.
\citeauthor{Narayanan2009} also found that the neural activity
in the mPFC changed significantly depending on
the outcome of the previous trial;
the first two principal components
change significantly after errors,
and appear to contain information that an error occurred.
This information would be necessary
in order to adapt behavior based on the last trial.

Interestingly, in the post-correct case,
the first principal component
is highly correlated with the integral
(i.e., cumulative sum) of the second principal component
during the post-correct trial
(Figure~\ref{fig2}E).
Since the first two principal components
account for over half of the variance
in the neural data (Figure~\ref{fig2}B),
we hypothesize that the mPFC
represents the task state (PC1)
and the integral of the task state (PC2).
Implementing integration
in a population of neurons has
been widely explored
\cite{Seung1996,Shadlen2001,Wang2002,Eliasmith2005}.
In the post-error case,
these correlations are significantly weaker,
indicating that the integration
taking place in the mPFC
depends on prior outcomes.
The firing patterns of individual neurons,
related to the two leading PCs,
are shown for post-correct trials in Figure~\ref{fig2}F.
Nearly equal fractions of neurons
became more or less active at times
when the PCs were maximally positive or negative.
In these plots, neural activity
is synchronized to the start of the trial (time 0).
The cue was presented at 1 second.
The animals were required to respond
to the cue with a reaction time less than 0.6 second,
and feedback about the outcome
was given approximately 0.1 second later.
The range of feedback times are indicated
near the upper right corner of each plot.
Importantly, the trials used for this analysis,
and shown in Figure~\ref{fig2}, were all correct trials.
The only difference between trial types
was the outcome of the previous trial
(correct or incorrect).
As in \citetext{Narayanan2008},
neural activity did not track
the type of error that occurred
(premature or late).
Population activity was not analyzed
for these subtypes of errors,
as they were present in unequal fractions of trials
over rats and were much less frequent
than the previously correct trials.

@@

\begin{table}
  \begin{center}
    \begin{tabular}{ccccc}
      \toprule
      $\beta$ & $R$ & $E$ \\
      \midrule
      0.44 & 2 & 0.5 \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Parameters used to simulate the dynamical system
    described in Equation~\eqref{eq:dint}.
    Appropriate values were found through simulation;
    see Figure~\ref{fig4}.}
  \label{tab:sim-param}
\end{table}

The purpose of the double integrator
is to predict the time at which the cue occurs.
Figure~\ref{fig4} shows that this is possible.
The parameter $\beta$ controls how fast
the second dimension of the double integrator state
($x_2$) increases (see Equation~\eqref{eq:dint}),
resulting in different states at $t_c$
(see Figure~\ref{fig4}B);
since this parameter is encoded in the strength
of the connection between the two integrators,
this is a natural target for learning
after errors.
Given a particular value for $\beta$,
the state of the system at the time of the cue,
$t_c$, can be predicted given
the initial conditions of the double integrator
at the start of the trial;
values at $t_c$ for $\beta=0.44$ are shown
in Figure~\ref{fig4}A.

The state of the system at the start of a trial
(i.e., the initial conditions of the dynamical system)
depend on the outcome of the previous trial.
Equation \eqref{eq:dint} describes how the state changes
based on the two possible outcomes,
reward or light extinguishing.
For sufficiently high values of $R$
(see Figure~\ref{fig4}C),
the system starts the next trial
near the origin,
resulting in the system predicting
the time of the cue,
and therefore employing the adaptive strategy.
For sufficiently high values of $E$
(see Figure~\ref{fig4}D),
the system starts the next trial near $(-1, -1)$,
resulting in the system being unable to predict
the time of the cue (see Figure~\ref{fig4}A),
and therefore employing the cue-responding strategy.
Note that correct and error trials
can be distinguished,
and therefore used to modify behavior
and drive learning;
however, premature trials
and late trials cannot be distinguished
on the next trial,
which is consistent with the results of
\citetext{Narayanan2009}.

@@

Equation~\eqref{eq:release} defines a signal
that causes the network to release the lever
when the predicted cue approaches.
Figure~\ref{fig5} shows a representative
correct trial from the cue-responding
and adaptive control networks,
as well as a trial in which
the cue was mispredicted,
and the subsequent corrected trials
following the error.
Importantly, Figure~\ref{fig5}B
shows the adaptive control network
can release the lever
at the exact time of the cue.

@@

The state of the system reflects the outcome
of the previous trial, and therefore
can affect behavior on the next trial.
The system can also be used to drive
learning in order to maximize long-term reward.
While the proposed system
does not employ learning,
previous work has proposed that the
mPFC drives learning of action outcomes.
Specifically, \citetext{Alexander2011} have proposed
the PRO model, in which the mPFC tracks the predicted values
of action outcomes in order to compute
prediction errors that drive learning.
While the signals tracked
by the double integrator model
exist to modify action timing,
the signals used to drive learning
in PRO can be computed from
the double integrator model as well,
as shown in Figure~\ref{fig6}.
The only difference that cannot
be attributed to the task
is that the state is tracked throughout
the trial in the PRO model,
while in the double integrator,
reward turns off the integrator
responsible for tracking task state.
Despite this difference,
the similarity in signals
tracked by the two models
suggests that the double integrator,
which we use to adaptively control action timing,
can also be used to drive learning.

@@

Figure~\ref{fig7} compares
the two leading principal components
of the spiking implementation of the double integrator network
to those of the experimental data.
The simulated principal components
are very similar to the experimental principal components
\cite{Narayanan2009}:
Pearson $R^2 = 0.82, 0.97, 0.95$
for the first component
and $R^2 = 0.91, 0.81, 0.84$
for the second components in
the post-correct, post-premature,
and post-late trials respectively.

@@

We also performed principal component analysis
on neural populations representing only one
of the two dimensions tracked by
the double integrator network.
Figure~\ref{fig8}
shows the results of the analysis
for the population tracking $x_1$.
While the first principal component
closely matches the experimental data
(Pearson $R^2 = 0.54, 0.86, 0.80$
for post-correct, post-premature, and post-late
trials respectively),
the second principal component
does not (Pearson $R^2 = 0.15, 0.26, 0.30$).
This suggests that there exist
neurons that are sensitive
to both dimensions of the double integrator.

@@

Figure~\ref{fig9} shows that
the decoded values of spiking neural networks
implementing the cue-responding
and adaptive control networks
closely resemble the signals
directly simulated with
the dynamical system
in correct, premature, and late trials.
Despite the introduction of the oscillation
and the additional injected noise current,
the cue can be predicted,
and the outcome of the previous trial
has a significant effect
on the system state during a trial.
The spiking networks are capable
of the adaptive control described
in the theoretical results,
as can be seen in Figure~\ref{fig9}A.

@@

Figure~\ref{fig10} shows the performance
of the experimental subjects,
the spiking cue-responding network,
and the full spiking adaptive control network.
The lengths of the simulations were
varied to match the number of trials
performed by subjects.
The mean of the median reaction times
is 272$\pm$48ms for the experimental subjects,
365$\pm$24ms for the cue-responding models,
and 240$\pm$62ms for the adaptive control model.
The mean performance (\% correct trials) is
70.8\% for experimental subjects,
98.6\% for cue-responding models,
and 92.5\% for adaptive control models.
Therefore, only the adaptive control model
has statistically indistinguishable reaction times
compared to the experimental subjects.
The cue-responding network is significantly slower,
but makes fewer errors than the adaptive control network.
Both simulated models make fewer errors
than experimental subjects.

@@

An interesting feature of the model
is that there is significant randomness
involved in model generation (see section~\ref{sec:nef}).
This results in significant variability
in the performance of simulated subjects,
despite all but one performing the task well
(see Figure~\ref{fig10}).
Also notable is that the model
does not employ learning:
the weights determined
before the simulation do not change.
The ability to keep track of the state
of the system over time through recurrent connections
is what enables the model to adapt.
By defining the dynamics of other tasks
and implementing them
in recurrently connected neural networks,
we can start to form general theories
of how biological systems can learn
to perform complex tasks.

@@

In this study, we make
an adaptive control hypothesis
of mPFC function in reaction time tasks.
This hypothesis is compatible
with previous hypotheses suggesting
that the mPFC plays a role
in the regulation of
reinforcement learning parameters.
One such example is the PRO
model by \citetext{Alexander2011}
(though these arguments also
apply to related models,
such as \citenoparens{Khamassi2012}).
The PRO model proposes that the mPFC
is involved in a form of
temporal difference learning.
As shown in Figure~\ref{fig6},
the positive and negative reward prediction error
computed by PRO can also be computed
by the double integrator,
assuming that the value function
learned by the system steadily increases.

@@

We have shown that neurons
in mPFC are tuned to multiple aspects
of the simple RT task
(see Figures~\ref{fig7} and \ref{fig8}).
Since we find neurons that are sensitive
to both task state ($x_1$) and the
length of time in that state ($x_2$),
we predict that single neurons
would be sensitive to additional
aspects of the task in more complex tasks.
This is due to the mPFC as a whole
tracking the dynamics of that task
in order to modify behavior.
Initial support for this idea
can be found in a recent study
of the mPFC that used
a delayed spatial alternation task
and found similar dynamics
in principal component space across
the mPFC ensembles \cite{Horst2012}.
This prediction could be further tested
by using a task that
has different temporal dynamics
in different trial blocks or task contexts,
and determining if there are neurons
with loadings on many primary components
in a principal component analysis.
One such task was used by
\citetext{Delatour2001}.

@@

\clearpage
\section{Figures}

\scalefigone{fig1}{1.0}{}

\scalefigone{fig2}{1.0}{}

\scalefigone{fig3}{0.8}{}

\scalefigone{fig4}{1.0}{}

\scalefigone{fig5}{1.0}{}

\scalefigone{fig6}{0.5}{}

\scalefigone{fig7}{1.0}{}

\scalefigone{fig8}{1.0}{}

\scalefigone{fig9}{0.5}{}

\scalefigone{fig10}{1.0}{}

>>>>

@@

\textbf{Pages:} 27. \textbf{Figures:} 10. \textbf{Tables:} 2. \textbf{Multimedia, 3D models:} 0

@@

The experimental data analyzed in this study is the same
as that in \citetext{Narayanan2009},
where more detailed explanation of the experimental
methods can be found.
In that study, twelve male Long-Evans rats
were trained to press down a lever
for a fixed amount of time (the ``foreperiod'';
1 second for this study)
using standard methods \cite{Laubach2000}.
After the foreperiod, an auditory cue is presented.
The subject then has 0.6 seconds (the response window)
in which to release the lever.
If the lever is released in that time,
the trial is a success, and the subject is rewarded
with water.
If the lever is released during the foreperiod,
the trial is classified as a ``premature error,''
and is penalized by a time-out with the house lights extinguished.
If the lever is not released within the response window,
the trial is classified as a ``late error,''
and is also penalized by a time-out with the house lights extinguished.
Figure~1A presents a schematic view of the task.

@@

In order to provide appropriate
environmental input for
the networks described
in subsequent sections,
the same reaction time task is simulated
using a finite state machine.
The task can be in one of seven states~($S$):
trial start~(TS), foreperiod~(FP),
timeout~(TO), cue~(C), response window~(R),
reward~(Rw) and intertrial interval~(ITI).
State transitions can be seen in
the graphical depiction of
the finite state machine in Figure~1C.

@@

In \citetext{Singh2006}, neural recordings
were analyzed during a single trial
of the vibrotactile discrimination task.
This corresponds to a single trial
of the simple RT task
in which the input, $u$,
is the lever press,
and the initial state, $x$, is $(0, 0)$.
However, in the simple RT task we perform
many consecutive trials.
We therefore use other inputs from
the environment to match
the trajectories tracked by
the experimental principal components
(shown in Figure~2).

Since the principal components
resemble the double integrator network
in the post-correct case,
we aim to reset the system to the origin point
after correct trials.
Figure~2D indicates that
the principal components return to their
original state after approximately 2 seconds,
which is the time at which
the subject receives reward.
Therefore, we hypothesize that
the trial's outcome (i.e., reward delivery
or a timeout) resets the system
to some particular state,
as proposed in \citetext{Narayanan2009}.
In the case of reward delivery,
we implement this hypothesis
by using the reward delivery signal, $u_{Rw}$,
to drive the system toward $(0, 0)$.
\begin{align*}
  \dot{x}_1 &= u - R \, u_{Rw} \, x_1 \\
  \dot{x}_2 &= \beta x_1 - R \, u_{Rw} \, x_2,
\end{align*}
where $R$ is a scalar parameter
indicating how strongly the system
is driven to the origin point.

In the post-error case,
the principal components
in Figure~2D are
not at the origin point.
Since the lever press increases $x_1$,
when $x_1 > 0$ we can conclude that
the subject is currently performing the task.
We therefore hypothesize that $x_1 < 0$
indicates that an error has recently occurred.
We implement this hypothesis
by using the house lights signal,
$u_{TO}$, to decrease $x_1$.
\begin{equation} \label{eq:dint}
  \begin{aligned}
    \dot{x}_1 &= u - R \, u_{Rw} \, x_1 - E u_{TO} \\
    \dot{x}_2 &= \beta x_1 - R \, u_{Rw} \, x_2,
  \end{aligned}
\end{equation}
where $E$ is a scalar parameter
controlling how quickly $x_1$ decreases.
Note that $x_2$ is not directly driven by $u_{TO}$;
however, because of the $\beta x_1$ term,
$x_2$ will also decrease.

Finally, Figure~2D shows that
the principal components are unstable
during the intertrial interval.
To model trial-by-trial cycles of
excitability associated with task engagement,
a slow oscillation ($\sim0.2$ Hz)
was applied to the networks.
The final dynamics of the double integrator are
\begin{equation} \label{eq:dint-o}
  \begin{aligned}
    \dot{x}_1 &= u - R \, u_{Rw} \, x_1 - E u_{TO} + \sin(0.2 \cdot 2 \pi) \\
    \dot{x}_2 &= \beta x_1 - R \, u_{Rw} \, x_2,
  \end{aligned}
\end{equation}
where $\beta$, $R$, and $E$ are positive scalars.
The effects of changing these free parameters
are explored in Section~\ref{sec:params}.

@@

The previous sections describe
a dynamical system hypothesis
of the mPFC and surrounding areas
of cortex that together
perform the simple reaction time task
adaptively depending on the outcome
of the previous trial.
While these can be numerically simulated
to observe ideal system behavior,
we construct a spiking neural network
implementing the above equations
in order to test the hypothesis
in a biologically plausible setting
that approximates the direct numerical simulation.
Testing this neural implementation
allows us to determine if the approximation
is able to perform the task
in a manner comparable to the real brain
by producing simulated neural data
that can be compared directly with
experimental neural data.
We construct spiking neural networks
using the principles of the Neural Engineering Framework
(NEF; \citenoparens{Eliasmith2003})
in the Nengo simulation environment
\cite{BekolayInpress}.
Using the NEF, each signal
from the dynamical systems above
are represented with a population of spiking neurons,
and transformed through the connections
between those populations
(see Figure~3).

@@

Simulated spike trains were generated
by the Nengo simulator.
The populations representing signals
in the double integrator and control
networks (see Sections~\ref{sec:dint} and \ref{sec:control})
contained 1200 adaptive leaky integrate-and-fire neurons
(see Figure~3);
174 of the neurons in an analyzed population
were randomly selected
in order to match the number of units
from the experimental study.

@@

As previously described in \citetext{Narayanan2009}
and summarized in Figure~2,
the first two principal components
of the experimental data
are modulated around the lever press
and appear to track
whether the subject has pressed the lever and
the relative amount of time
that the lever has been pressed.
\citeauthor{Narayanan2009} also found that the neural activity
in the mPFC changed significantly depending on
the outcome of the previous trial;
the first two principal components
change significantly after errors,
and appear to contain information that an error occurred.
This information would be necessary
in order to adapt behavior based on the last trial.

Interestingly, in the post-correct case,
the first principal component
is highly correlated with the integral
(i.e., cumulative sum) of the second principal component
during the post-correct trial
(Figure~2E).
Since the first two principal components
account for over half of the variance
in the neural data (Figure~2B),
we hypothesize that the mPFC
represents the task state (PC1)
and the integral of the task state (PC2).
Implementing integration
in a population of neurons has
been widely explored
\cite{Seung1996,Shadlen2001,Wang2002,Eliasmith2005}.
In the post-error case,
these correlations are significantly weaker,
indicating that the integration
taking place in the mPFC
depends on prior outcomes.
The firing patterns of individual neurons,
related to the two leading PCs,
are shown for post-correct trials in Figure~2F.
Nearly equal fractions of neurons
became more or less active at times
when the PCs were maximally positive or negative.
In these plots, neural activity
is synchronized to the start of the trial (time 0).
The cue was presented at 1 second.
The animals were required to respond
to the cue with a reaction time less than 0.6 second,
and feedback about the outcome
was given approximately 0.1 second later.
The range of feedback times are indicated
near the upper right corner of each plot.
Importantly, the trials used for this analysis,
and shown in Figure~2, were all correct trials.
The only difference between trial types
was the outcome of the previous trial
(correct or incorrect).
As in \citetext{Narayanan2008},
neural activity did not track
the type of error that occurred
(premature or late).
Population activity was not analyzed
for these subtypes of errors,
as they were present in unequal fractions of trials
over rats and were much less frequent
than the previously correct trials.

@@

\begin{table}
  \begin{center}
    \begin{tabular}{ccccc}
      \toprule
      $\beta$ & $R$ & $E$ \\
      \midrule
      0.44 & 2 & 0.5 \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Parameters used to simulate the dynamical system
    described in Equation~\eqref{eq:dint}.
    Appropriate values were found through simulation;
    see Figure~4.}
  \label{tab:sim-param}
\end{table}

The purpose of the double integrator
is to predict the time at which the cue occurs.
Figure~4 shows that this is possible.
The parameter $\beta$ controls how fast
the second dimension of the double integrator state
($x_2$) increases (see Equation~\eqref{eq:dint}),
resulting in different states at $t_c$
(see Figure~4B);
since this parameter is encoded in the strength
of the connection between the two integrators,
this is a natural target for learning
after errors.
Given a particular value for $\beta$,
the state of the system at the time of the cue,
$t_c$, can be predicted given
the initial conditions of the double integrator
at the start of the trial;
values at $t_c$ for $\beta=0.44$ are shown
in Figure~4A.

The state of the system at the start of a trial
(i.e., the initial conditions of the dynamical system)
depend on the outcome of the previous trial.
Equation \eqref{eq:dint} describes how the state changes
based on the two possible outcomes,
reward or light extinguishing.
For sufficiently high values of $R$
(see Figure~4C),
the system starts the next trial
near the origin,
resulting in the system predicting
the time of the cue,
and therefore employing the adaptive strategy.
For sufficiently high values of $E$
(see Figure~4D),
the system starts the next trial near $(-1, -1)$,
resulting in the system being unable to predict
the time of the cue (see Figure~4A),
and therefore employing the cue-responding strategy.
Note that correct and error trials
can be distinguished,
and therefore used to modify behavior
and drive learning;
however, premature trials
and late trials cannot be distinguished
on the next trial,
which is consistent with the results of
\citetext{Narayanan2009}.

@@

Equation~\eqref{eq:release} defines a signal
that causes the network to release the lever
when the predicted cue approaches.
Figure~5 shows a representative
correct trial from the cue-responding
and adaptive control networks,
as well as a trial in which
the cue was mispredicted,
and the subsequent corrected trials
following the error.
Importantly, Figure~5B
shows the adaptive control network
can release the lever
at the exact time of the cue.

@@

The state of the system reflects the outcome
of the previous trial, and therefore
can affect behavior on the next trial.
The system can also be used to drive
learning in order to maximize long-term reward.
While the proposed system
does not employ learning,
previous work has proposed that the
mPFC drives learning of action outcomes.
Specifically, \citetext{Alexander2011} have proposed
the PRO model, in which the mPFC tracks the predicted values
of action outcomes in order to compute
prediction errors that drive learning.
While the signals tracked
by the double integrator model
exist to modify action timing,
the signals used to drive learning
in PRO can be computed from
the double integrator model as well,
as shown in Figure~6.
The only difference that cannot
be attributed to the task
is that the state is tracked throughout
the trial in the PRO model,
while in the double integrator,
reward turns off the integrator
responsible for tracking task state.
Despite this difference,
the similarity in signals
tracked by the two models
suggests that the double integrator,
which we use to adaptively control action timing,
can also be used to drive learning.

@@

Figure~7 compares
the two leading principal components
of the spiking implementation of the double integrator network
to those of the experimental data.
The simulated principal components
are very similar to the experimental principal components
\cite{Narayanan2009}:
Pearson $R^2 = 0.82, 0.97, 0.95$
for the first component
and $R^2 = 0.91, 0.81, 0.84$
for the second components in
the post-correct, post-premature,
and post-late trials respectively.

@@

We also performed principal component analysis
on neural populations representing only one
of the two dimensions tracked by
the double integrator network.
Figure~8
shows the results of the analysis
for the population tracking $x_1$.
While the first principal component
closely matches the experimental data
(Pearson $R^2 = 0.54, 0.86, 0.80$
for post-correct, post-premature, and post-late
trials respectively),
the second principal component
does not (Pearson $R^2 = 0.15, 0.26, 0.30$).
This suggests that there exist
neurons that are sensitive
to both dimensions of the double integrator.

@@

Figure~9 shows that
the decoded values of spiking neural networks
implementing the cue-responding
and adaptive control networks
closely resemble the signals
directly simulated with
the dynamical system
in correct, premature, and late trials.
Despite the introduction of the oscillation
and the additional injected noise current,
the cue can be predicted,
and the outcome of the previous trial
has a significant effect
on the system state during a trial.
The spiking networks are capable
of the adaptive control described
in the theoretical results,
as can be seen in Figure~9A.

@@

Figure~10 shows the performance
of the experimental subjects,
the spiking cue-responding network,
and the full spiking adaptive control network.
The lengths of the simulations were
varied to match the number of trials
performed by subjects.
The mean of the median reaction times
is 272$\pm$48ms for the experimental subjects,
365$\pm$24ms for the cue-responding models,
and 240$\pm$62ms for the adaptive control model.
The mean performance (\% correct trials) is
70.8\% for experimental subjects,
98.6\% for cue-responding models,
and 92.5\% for adaptive control models.
Therefore, only the adaptive control model
has statistically indistinguishable reaction times
compared to the experimental subjects.
The cue-responding network is significantly slower,
but makes fewer errors than the adaptive control network.
Both simulated models make fewer errors
than experimental subjects.

@@

An interesting feature of the model
is that there is significant randomness
involved in model generation (see section~\ref{sec:nef}).
This results in significant variability
in the performance of simulated subjects,
despite all but one performing the task well
(see Figure~10).
Also notable is that the model
does not employ learning:
the weights determined
before the simulation do not change.
The ability to keep track of the state
of the system over time through recurrent connections
is what enables the model to adapt.
By defining the dynamics of other tasks
and implementing them
in recurrently connected neural networks,
we can start to form general theories
of how biological systems can learn
to perform complex tasks.

@@

In this study, we make
an adaptive control hypothesis
of mPFC function in reaction time tasks.
This hypothesis is compatible
with previous hypotheses suggesting
that the mPFC plays a role
in the regulation of
reinforcement learning parameters.
One such example is the PRO
model by \citetext{Alexander2011}
(though these arguments also
apply to related models,
such as \citenoparens{Khamassi2012}).
The PRO model proposes that the mPFC
is involved in a form of
temporal difference learning.
As shown in Figure~6,
the positive and negative reward prediction error
computed by PRO can also be computed
by the double integrator,
assuming that the value function
learned by the system steadily increases.

@@

We have shown that neurons
in mPFC are tuned to multiple aspects
of the simple RT task
(see Figures~7 and 8).
Since we find neurons that are sensitive
to both task state ($x_1$) and the
length of time in that state ($x_2$),
we predict that single neurons
would be sensitive to additional
aspects of the task in more complex tasks.
This is due to the mPFC as a whole
tracking the dynamics of that task
in order to modify behavior.
Initial support for this idea
can be found in a recent study
of the mPFC that used
a delayed spatial alternation task
and found similar dynamics
in principal component space across
the mPFC ensembles \cite{Horst2012}.
This prediction could be further tested
by using a task that
has different temporal dynamics
in different trial blocks or task contexts,
and determining if there are neurons
with loadings on many primary components
in a principal component analysis.
One such task was used by
\citetext{Delatour2001}.

@@
